---
title: "Assignment 3"
author: "Student ID: 201903536"
date: "30 November 2023"
output: html_document
---

```{r setup, include=FALSE} 
#####################################
# SETUP
#####################################

knitr::opts_chunk$set(echo = FALSE) 

#####################################
# Install/load packages
#####################################


```


## GitHub

The GitHub repository for **this assignment** can be found [here](https://github.com/emmi3105/472_assignment3).


## Exercise 1


## Exercise 2

Write an automatic webscraping function in R that constructs a table (as e.g. a data frame, a tibble, or a data table) of all R1 (Very High Research Activity) and R2 (High Research Activity) Research Universities in the United States of America. These data can be found on wikipedia.

Your initial scraper should collect five variables:

I. The university’s name
It’s status (public or private)
The city in which it is located
The state in which it is located
The URL of the university’s dedicated Wikipedia page


## Exercise 3



## Data



## Sources



## Appendix: All code in this assignment

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE} 


```
